Finch: Sparse and Structured Array Programming with Control Flow

### 2024-04-01 sections/tensors.tex
comment

To build our bridge, we embrace a set of abstractions: level formats/Fiber Trees, iteration context dependent instantiation of iterators, and tensor life cycles.
%
Our first abstraction mostly already exists in the literature: a manner of specifying a data structure for a multi-dimensional tensor out of data structures for single dimensional tensors~\cite{sze2017efficient,chou2022compilation, chou2018format}.
%
We recapitulate the essential details here.
%
Our next two abstractions add to to the first by providing a mechanism to use data structures generated by the first abstraction in a greater variety of contexts while maintaining per-dimension encapsulation of array data structures.
%
We introduce an interface to instatiate iterators in a variety of contexts in our programs and we introduce the lifecycle interface to manage when we read and write to multi-dimensional iterators.
%
These interfaces add to the level abstraction, expanding the types of data that they can express via mapping to looplets and expanding the contexts in which they can be used.
%
Previous efforts to compile a greater variety of sparse array programs left these bridges untouched ~\cite{henry_compilation_2021, won2023unified, senanayake2020sparse}.

>>>>

\willow{After making changes to the introduction of this section, some of the following text feels redundant}
To build our bridge, we embrace a set of abstractions: level formats/Fiber Trees, iteration context dependent instantiation of iterators, and tensor life cycles.
%
Our first abstraction mostly already exists in the literature: a manner of specifying a data structure for a multi-dimensional tensor out of data structures for single dimensional tensors~\cite{sze2017efficient,chou2022compilation, chou2018format}.
%
We recapitulate the essential details here.
%
Our next two abstractions add to to the first by providing a mechanism to use data structures generated by the first abstraction in a greater variety of contexts while maintaining per-dimension encapsulation of array data structures.
%
We introduce an interface to instatiate iterators in a variety of contexts in our programs and we introduce the lifecycle interface to manage when we read and write to multi-dimensional iterators.
%
These interfaces add to the level abstraction, expanding the types of data that they can express via mapping to looplets and expanding the contexts in which they can be used.
%
Previous efforts to compile a greater variety of sparse array programs left these bridges untouched ~\cite{henry_compilation_2021, won2023unified, senanayake2020sparse}.
